Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-27 WSSSPE1_report/WSSSPE1_summary_report.tex
more reading and small changes

Papers and discussion concentrated in three areas: How then ought software work to be visible, what are the barriers to its visibility, and what can be done to make it more visible?

@@

Another approach is to try to reduce the difficulty of citing software for authors, acknowledging that authors are often working with software that itself wraps other software and therefore hides that software. Knepley et al approach this by proposing a mechanism by which the software itself, after it has run, provides the user with a set of citations that are relevant to the code actually run. They describe a prototype implementation whereby the citations are embedded in libraries and reported along with the results, via a commandline interface~\cite{Knepley_WSSSPE}. Discussion highlighted the difficulty that attempting to acknowledge the contributions of all pieces of dependent code within a paper faces the difficulty of creating very long citation lists, straining the analogy of code used to papers cited. Katz approaches this issue by proposing a system of ``transitive credit,'' recording dependencies and relative contributions outside particular papers, relieving authors from the responsibility of acknowledging each and every dependency. In stead authors would acknowledge the percentage contribution of the software they used directly and an external system would then be able to recursively allocate that credit to those who had provided dependencies. Finally Priem and Piwowar argued that machine learning techniques could examine the body of published literature and extract mentions of software, coping with the multitude of informal ways in which authors mention software they have used~\cite{Priem_WSSSPE}.  Discussion included turning the emphasis from asking users to improve their citation of software contributions, to ask how projects producing software might monitor the literature to improve their ability to show impact. Michael McLennan described the approach taken by the NanoHub project to scan the literature using keywords and names of known users to discover papers that are likely to have used their software and platform, then to have graduate students read each paper, highlighting any mention and perhaps following up with the authors to locate stories of impact.  Work since the workshop has described this process at publications.wikia.com.

Acknowledgement in science does not come only in publications, of course. A key location for visibility might be in the grant funding process, both in bio-sketches and in grant project reporting. Some progress has been made here. Representatives of the NSF at the meeting emphasized these opportunities and encouraged participants to take advantage of them. Nonetheless there was skepticism that peer review panels would value these contributions in the same was as publications).

@@

The paper by Maheshwari et al.~\cite{Maheshwari_WSSSPE} focuses on ``technology
catalysts" and their role in the modern scientific research process. Technology
catalyst could be defined as a role played by an individual with a know-how of
technological advancements, tasked with user engagement with a goal of enacting
scientific or engineering applications, and using suitable tools and techniques
to take advantage of technological capabilities and benefiting applications.

One of the tasks of catalysts is to seek community collaborations for new
applications and user engagement thus benefiting both: science, by effective
running of scientific codes on computational infrastructure, and technology,
by conducting research and seeking findings for technology improvement.
The particular engagements described in the paper came up from authors work as
postdoctoral researcher at Cornell and Argonne. Interaction with the scientific
communities in both institutions resulted in these collaborations.

@@

Christopherson et al.~\cite{Christopherson_WSSSPE} outlines the degree to which research relies on high
quality software. There are often barriers and a lack of suitable incentives
for researchers to embrace software engineering principals. The Water Science
Software Institute is working to lower some of these barriers through an Open
Community Engagement Process. This is a 4-step iterative development process
that incorporates Agile development principals.

@@

One key to SciGaP is its openness. While based on the Apache Airavata project
and the CIPRES Workbench Framework, community contributions are encouraged
because of its open source, open governance and open operations policies. The
goal is robust, sustainable infrastructure with a cycle of development that
improves reliability and prioritizes stakeholder requirements. The project is
leveraging Internet2's Net+ mechanisms for converting SciGaP and its gateways
into commodity services.

>>>>

Papers and discussion concentrated in three areas: How then \note{something missing here in this sentence --ketan} ought software  work to be visible, what are the barriers to its visibility, and what can be done to make it more visible?

@@

Another approach is to try to reduce the difficulty of citing software for authors, acknowledging that authors are often working with software that itself wraps other software and therefore hides that software. Knepley et. al. approach this by proposing a mechanism by which the software itself, after it has run, provides the user with a set of citations that are relevant to the code actually run. They describe a prototype implementation whereby the citations are embedded in libraries and reported along with the results, via a commandline interface~\cite{Knepley_WSSSPE}. Discussion highlighted the difficulty that attempting to acknowledge the contributions of all pieces of dependent code within a paper faces the difficulty of creating very long citation lists, straining the analogy of code used to papers cited. Katz approaches this issue by proposing a system of ``transitive credit,'' recording dependencies and relative contributions outside particular papers, relieving authors from the responsibility of acknowledging each and every dependency. Instead authors would acknowledge the percentage contribution of the software they used directly and an external system would then be able to recursively allocate that credit to those who had provided dependencies. Finally Priem and Piwowar argued that machine learning techniques could examine the body of published literature and extract mentions of software, coping with the multitude of informal ways in which authors mention software they have used~\cite{Priem_WSSSPE}.  Discussion included turning the emphasis from asking users to improve their citation of software contributions, to ask how projects producing software might monitor the literature to improve their ability to show impact. Michael McLennan described the approach taken by the NanoHub project to scan the literature using keywords and names of known users to discover papers that are likely to have used their software and platform, then to have graduate students read each paper, highlighting any mention and perhaps following up with the authors to locate stories of impact.  Work since the workshop has described this process at \textit{publications.wikia.com}.

Acknowledgement in science does not come only in publications, of course. A key location for visibility might be in the grant funding process, both in bio-sketches and in grant project reporting. Some progress has been made here. Representatives of the NSF at the meeting emphasized these opportunities and encouraged participants to take advantage of them. Nonetheless, there was skepticism that peer review panels would value these contributions in the same ways as publications).

@@

The paper by Maheshwari et al.~\cite{Maheshwari_WSSSPE} focuses on ``technology
catalysts" and their role in the modern scientific research process. A technology
catalyst can be defined as a role played by an individual with a know-how of
technological advancements, tasked with user-engagement with a goal of enacting
scientific or engineering applications, and using suitable tools and techniques
to take advantage of technological capabilities thus benefiting applications.

One of the tasks of technology catalysts is to seek community collaborations for new
applications and user engagement thus benefiting both: science, by effective
running of scientific codes on computational infrastructure, and technology,
by conducting research and seeking findings for technology improvement.
The particular engagements described in the paper came up from authors work as
postdoctoral researcher at Cornell and Argonne. Interaction with the scientific
communities in both institutions resulted in these collaborations.

@@

Christopherson et al.~\cite{Christopherson_WSSSPE} outlines the degree to which research relies on high
quality software. There are often barriers and a lack of suitable incentives
for researchers to embrace software engineering principles. The Water Science
Software Institute is working to lower some of these barriers through an Open
Community Engagement Process. This is a 4-step iterative development process
that incorporates Agile development principles.

@@

One key to Sci-GaP is its openness. While based on the Apache Airavata project
and the CIPRES Workbench Framework, community contributions are encouraged
because of its open source, open governance and open operations policies. The
goal is robust, sustainable infrastructure with a cycle of development that
improves reliability and prioritizes stakeholder requirements. The project is
leveraging Internet2's Net+ mechanisms for converting SciGaP and its gateways
into commodity services.
